Pattern matching for\\ $(231,213)$-avoiding permutations

### 2015-01-31 article.tex
introduction

Suppoe $P$ is a set of permutations. We let $\AV_n(p)$ denote the
set of all $n$-permutations avoiding each permutation in $P$.
For the sake of convenience, we omit $P$'s braces thus having
e.g. $\AV_3(132,4312)$ instead of
$\AV_3(\{132,4312\})$.
A basic example is if 
$\pi = \pi_1\,\pi_2\,\ldots\,\pi_n$ avoids $3\,2\,1$, i.e., 
has no decreasing subsequence of length $3$, then its reverse, 
$\pi' = \pi_n\,\pi_{n-1}\,\ldots\,\pi_1$ avoids $1\,2\,3$, i.e., 
has no increasing subsequence of length $3$.

@@

		A \textbf{Stripe} of a permutation
		is a mapping into \textbf{up step} (noted $\ustep$) 
		or a \textbf{down step} (noted $\dstep$).
		An element $i$ of a permutation $p$ is an up step if and only if $i$ is not the last element and $p(i)<p(i+1)$.
		An element $i$ of a permutation $p$ is a down step if and only if $i$ is not the last element and $p(i)>p(i+1)$.\\	
		Formally a Stripe is a function $\mathcal{S} : [n-1] \to \{\dstep,\ustep\}$ define as :\\
		$ \mathcal{S}(p,i) = \left\{ 
		  \begin{array}{l l}
		    \ustep & \quad \text{if $1 \leq i<n$ and $p(i+1)-p(i)>0$ }\\
		    \dstep & \quad \text{if $1 \leq i<n$ and $p(i+1)-p(i)<0$ }\\
		  \end{array} \right. $\\

		By extension we define  $\mathcal{S} : \Perm{n} \to \{\dstep,\ustep\}^{n}$ such as :
		$\mathcal{S}(p)=\mathcal{S}(p,1)\mathcal{S}(p,2)...\mathcal{S}(p,n-1)$.
		We call $\mathcal{S}(p)$ the stripe of $p$.
		For example the stripe of the permutation $(12,11,10,1,2,3,4,9,8,5,6,7)$, 
		is $\dstep,\dstep,\dstep,\ustep,\ustep,\ustep,\ustep,\dstep,\dstep,\ustep,\ustep$.\\

%		\textbf{Corollary \num} The stripe function
%		for $\C{n}$ are in bijection with with $\{\ustep,\dstep\}^{n-1}$.\\

%		\textbf{Proposition \numl{updown}} Let $p \in \C{n}$,
%		$\forall x,y \in [n]$, 
%		$\mathcal{S}(p,x)=\ustep$
%		and $\\mathcal{S}(p,y)=\dstep$
%		then $p(x)<p(y)$.\\

		Formally we say that the stripe of $\pmotif \in \Perm{m}$ is a motif
		of a stripe of $\ptext \in \Perm{n}$ if and only if there exist an increasing function $\phi : [|\pmotif|] \rightarrow [|\pmotif|]$
		such that : \\
		$\forall i \in [|\pmotif|]$, $\mathcal{S}(\pmotif,i) = \mathcal{S}(\ptext,\phi(i))$\\
		If a stripe $R_2$ is a motif of a stripe $R_1$,
		it means that there exists
		a subsequence of $R_1$ such as
		this sequence is equal to $R_2$.
%		We say that a set is a match 
%		if and only if there exists a function such as define above
%		and the set is the set of image of this function.\\
		For example, the stripe $(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_1)$ is a motif of  the stripe \\
		$(\dstep_1,\dstep_2,\dstep_3,\ustep_1,\ustep_2,\ustep_3,\ustep_4,\dstep_4,\dstep_5,\ustep_5,\ustep_6)$ ,
		because we can match \\
		$(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_1)$ to $(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_5,\ustep_6)$.
		\begin{figure}[h!]
			\begin{displaymath}
				\xymatrix@=0.3cm{
					\dstep_1 \ar[dd] & \dstep_2 \ar[dd] & \dstep_3 \ar[dd] & \dstep_4\ar[ddrrrr]  & \ustep_1 \ar[ddrrrrr]\\
					\\
					\dstep_1 & \dstep_2 & \dstep_3 & \ustep_1 & \ustep_2 & \ustep_3 & \ustep_4 & \dstep_4 & \dstep_5 & \ustep_5 & \ustep_6 \\
				}
			\end{displaymath}     
		\end{figure}		
		But the stripe $(\dstep,\ustep,\dstep,\dstep,\dstep)$ is not a motif of the stripe\\ $(\dstep_1,\dstep_2,\dstep_3,\ustep_1,\ustep_2,\ustep_3,\ustep_4,\dstep_4,\dstep_5,\ustep_5,\ustep_6,\ustep_7)$.\\

		A \textbf{bivincular motif} generalize
		a motif. 
		A bivincular motif allows 
		to force element to be adjacent in position
		or/and in value.
		A bivincular motif
		is a motif with two set $\bmotif$,
		the first set contains the first index 
		of two consecutive indexes
		and the second set contains the first indexes
		of the two consecutive elements.
		Formally, Let $\bmotif$ 
		be a bivincular motif 
		and $\ptext$ a permutation.
		if $i \in \x$,
		and $i$ is matched with $j$ in $\ptext$
		then $i+1$ must be matched with $j+1$,
		with the special case of $0$ and $|\pmotif|$ :
		if $0 \in \x$ then the first element of $\pmotif$
		must be matched with the first element of $\ptext$,
		and if $|\pmotif| \in \x$ then the last element of $\pmotif$
		must be matched with the last element of $\ptext$.
		if $i \in \y$,
		and $i$ is matched with $j$ in $\ptext$
		then $i+1$ must be matched with $l$
		such as $\ptext(l)=\ptext(j)+1$,
		with the special case of $0$ and $|\pmotif|$ :
		if $0 \in \y$ then the minimal element of $\pmotif$
		must be matched with the minimal element of $\ptext$,
		and if $|\pmotif| \in \y$ then the maximal element of $\pmotif$
		must be matched with the maximal element of $\ptext$.

>>>>

Suppose $P$ is a set of permutations. We let $\AV_n(p)$ denote the
set of all $n$-permutations avoiding each permutation in $P$.
For the sake of convenience, we omit $P$'s braces thus having
e.g. $\AV_3(132,4312)$ instead of
$\AV_3(\{132,4312\})$.
A basic example is if 
$\pi = \pi_1\,\pi_2\,\ldots\,\pi_n$ avoids $3\,2\,1$, i.e., 
has no decreasing subsequence of length $3$, then its reverse, 
$\pi' = \pi_n\,\pi_{n-1}\,\ldots\,\pi_1$ avoids $1\,2\,3$, i.e., 
has no increasing subsequence of length $3$.

An \emph{ascent} of a permutation $\pi \in S_n$ is any position 
$1 \leq i < n$ where the following value is bigger than the current one. 
That is, if $\pi = \pi_1\,\pi_2\,\ldots\,\pi_n$, then
$i$ is an ascent if $\pi_i < \pi_{i+1}$.
For example, the permutation 
$3\,4\,5\,2\,1\,6\,$7 has ascents (at positions) $1,2,5,6$.
Similarly, a \emph{descent} is a position 
$1 \leq i < n$ with $\pi_i > \pi_{i+1}$, 
so every $i$ with $1 \leq i < n$ either is an ascent or is a descent of 
$\pi$.
Let $\ustep$ and $\dstep$ denote an ascend and a descend, respectively.
The \emph{stripe} $s_\pi$ of a permutation $\pi \in S_n$ is the word
$s_{\pi,1} s_{\pi,2} \ldots s_{\pi,n-1} \in \{\ustep,\dstep\}^{n-1}$ 
defined by 
$\mathcal{S}_{\pi,i} = \ustep$ if $i$ an ascent in $\pi$ and
$\mathcal{S}_\pi(i) = \dstep$ if $i$ a descent in $\pi$.
For example the stripe of the permutation 
$\pi = 12\,11\,10\,1\,2\,3,4\,9\,8\,5\,6\,7$ 
is 
$s_\pi = \dstep\dstep\dstep\ustep\ustep\ustep\ustep\dstep\dstep\ustep\ustep$.
The stripe $s_\sigma$ is a \emph{pattern} of the strip $s_\pi$
(or $s_\sigma$ \emph{occurs} of the strip $s_\pi$) if $s_\sigma$ occurs in
$s_\pi$ as a subsequence.

% For two permutations $\
% The stripe $s\pi$ of $\pi \in S-n$ is a motif pmotif \in \Perm{m}$ is a motif
% of a stripe of $\ptext \in \Perm{n}$ if and only if there exist an increasing function $\phi : [|\pmotif|] \rightarrow [|\pmotif|]$
% such that : \\
% $\forall i \in [|\pmotif|]$, $\mathcal{S}(\pmotif,i) = \mathcal{S}(\ptext,\phi(i))$\\
% If a stripe $R_2$ is a motif of a stripe $R_1$,
% it means that there exists
% a subsequence of $R_1$ such as
% this sequence is equal to $R_2$.

% For example, the stripe $(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_1)$ is a motif of  the stripe \\
% $(\dstep_1,\dstep_2,\dstep_3,\ustep_1,\ustep_2,\ustep_3,\ustep_4,\dstep_4,\dstep_5,\ustep_5,\ustep_6)$ ,
% because we can match \\
% $(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_1)$ to $(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_5,\ustep_6)$.
% \begin{figure}[h!]
% 	\begin{displaymath}
% 		\xymatrix@=0.3cm{
% 			\dstep_1 \ar[dd] & \dstep_2 \ar[dd] & \dstep_3 \ar[dd] & \dstep_4\ar[ddrrrr]  & \ustep_1 \ar[ddrrrrr]\\
% 			\\
% 			\dstep_1 & \dstep_2 & \dstep_3 & \ustep_1 & \ustep_2 & \ustep_3 & \ustep_4 & \dstep_4 & \dstep_5 & \ustep_5 & \ustep_6 \\
% 		}
% 	\end{displaymath}     
% \end{figure}		
% But the stripe $(\dstep,\ustep,\dstep,\dstep,\dstep)$ is not a motif of the stripe\\ $(\dstep_1,\dstep_2,\dstep_3,\ustep_1,\ustep_2,\ustep_3,\ustep_4,\dstep_4,\dstep_5,\ustep_5,\ustep_6,\ustep_7)$.\\

% A \textbf{bivincular motif} generalize
% a motif. 
% A bivincular motif allows 
% to force element to be adjacent in position
% or/and in value.
% A bivincular motif
% is a motif with two set $\bmotif$,
% the first set contains the first index 
% of two consecutive indexes
% and the second set contains the first indexes
% of the two consecutive elements.
% Formally, Let $\bmotif$ 
% be a bivincular motif 
% and $\ptext$ a permutation.
% if $i \in \x$,
% and $i$ is matched with $j$ in $\ptext$
% then $i+1$ must be matched with $j+1$,
% with the special case of $0$ and $|\pmotif|$ :
% if $0 \in \x$ then the first element of $\pmotif$
% must be matched with the first element of $\ptext$,
% and if $|\pmotif| \in \x$ then the last element of $\pmotif$
% must be matched with the last element of $\ptext$.
% if $i \in \y$,
% and $i$ is matched with $j$ in $\ptext$
% then $i+1$ must be matched with $l$
% such as $\ptext(l)=\ptext(j)+1$,
% with the special case of $0$ and $|\pmotif|$ :
% if $0 \in \y$ then the minimal element of $\pmotif$
% must be matched with the minimal element of $\ptext$,
% and if $|\pmotif| \in \y$ then the maximal element of $\pmotif$
% must be matched with the maximal element of $\ptext$.

@@
